\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta, positioning, fit, backgrounds}
\usepackage{listings}
\usepackage{natbib}
\usepackage{caption}

\hypersetup{
    colorlinks=true,
    linkcolor=blue!70!black,
    citecolor=blue!70!black,
    urlcolor=blue!70!black,
}

\lstset{
    basicstyle=\small\ttfamily,
    keywordstyle=\color{blue!70!black},
    commentstyle=\color{green!50!black},
    stringstyle=\color{red!70!black},
    frame=single,
    breaklines=true,
    columns=fullflexible,
}

\title{\textbf{RubinAlerts}: A Multi-Broker Pipeline for Type~Ia Supernova\\
Discovery and Spectroscopic Follow-Up Planning\\
in Rubin Observatory Deep Drilling Fields}

\author{C.~Stubbs}

\date{February 2026}

\begin{document}

\maketitle

\begin{abstract}
We describe \textsc{RubinAlerts}, a pipeline for discovering, aggregating,
and prioritizing Type~Ia supernova (SN~Ia) candidates in the Rubin
Observatory Legacy Survey of Space and Time (LSST) Deep Drilling Fields
(DDFs).  The system queries three independent alert brokers---ANTARES,
ALeRCE (ZTF), and ALeRCE-LSST---cross-matches their detections,
deduplicates sources, screens against known variable star catalogs, and
produces a ranked list of SN~Ia candidates.  Galactic extinction
corrections are applied using IRSA SFD dust maps, and spectroscopic
redshifts are obtained from NED for host galaxy distance estimation.
Light curve peak fitting is performed using both inverted-parabola
models in flux space and the Villar supernova parametric model (SPM),
with peak times determined by numerical flux maximization.  A
two-dimensional merit function combining temporal proximity to peak and
extinction-corrected apparent brightness enables automated generation of
prioritized spectroscopic follow-up target lists for the Magellan
telescopes at Las Campanas Observatory.  This document describes the
pipeline architecture, algorithms, data structures, and current
implementation status.
\end{abstract}

\tableofcontents
\newpage

%% =========================================================================
\section{Introduction}
\label{sec:intro}
%% =========================================================================

Type~Ia supernovae (SNe~Ia) serve as standardizable candles for
measuring cosmological distances.  The Vera C.\ Rubin Observatory's
LSST will discover thousands of SNe~Ia in its Deep Drilling Fields,
observed with a cadence of $\sim$3--5~days in multiple optical bands
($ugrizy$).  However, photometric discovery alone is insufficient for
precision cosmology: spectroscopic classification and redshift
measurement are essential, and telescope time on facilities such as the
Magellan 6.5\,m telescopes is limited.

The challenge is therefore one of \emph{triage}: given hundreds of
photometric SN candidates discovered each night, which ones should
receive scarce spectroscopic follow-up resources?  The answer depends on
(1)~the probability that the candidate is a genuine SN~Ia rather than a
contaminant, (2)~its brightness relative to the spectroscopic
sensitivity limit, and (3)~its temporal phase relative to maximum
light---spectra obtained near peak are most valuable for classification
and cosmological standardization.

\textsc{RubinAlerts} addresses this by aggregating classifications from
multiple alert brokers, fitting light curves to estimate peak
brightness and timing, and computing a merit function that encodes the
scientific value of observing each candidate on a given night.  The
output is a prioritized, observatory-ready target list in the native
Magellan catalog format.

\subsection{Deep Drilling Fields}

The pipeline targets the seven fields listed in
Table~\ref{tab:ddf_fields}, comprising the six Rubin LSST DDFs plus the
M49 (NGC~4472) field in the Virgo Cluster.  Each field is searched
within a 1.75\degr\ radius, matching the LSST field of view
($\sim$3.5\degr\ diameter).

\begin{table}[h]
\centering
\caption{Target fields and their properties.  The ``ZTF'' column
indicates whether the field falls within the Zwicky Transient Facility
footprint (Dec~$\gtrsim -30\degr$).}
\label{tab:ddf_fields}
\begin{tabular}{lrrcc}
\toprule
Field & RA (deg) & Dec (deg) & Search radius & ZTF coverage \\
\midrule
COSMOS   & 150.11 & $+2.23$  & 1.75\degr & Yes \\
XMM-LSS  &  35.57 & $-4.82$  & 1.75\degr & Yes \\
ECDFS    &  52.98 & $-28.12$ & 1.75\degr & Marginal \\
ELAIS-S1 &   9.45 & $-44.02$ & 1.75\degr & No \\
EDFS\_a  &  58.90 & $-49.32$ & 1.75\degr & No \\
EDFS\_b  &  63.60 & $-47.60$ & 1.75\degr & No \\
M49      & 187.44 & $+8.00$  & 1.75\degr & Yes \\
\bottomrule
\end{tabular}
\end{table}


%% =========================================================================
\section{Pipeline Architecture}
\label{sec:architecture}
%% =========================================================================

Figure~\ref{fig:pipeline} shows the end-to-end data flow.  The pipeline
is implemented in Python and orchestrated through a Jupyter notebook
(\texttt{supernova\_monitor.ipynb}), with reusable logic organized into
modules under the \texttt{core/}, \texttt{broker\_clients/},
\texttt{cache/}, \texttt{host\_galaxy/}, and \texttt{utils/}
directories.

\begin{figure}[h]
\centering
\begin{tikzpicture}[
    node distance=0.8cm and 1.2cm,
    block/.style={rectangle, draw, fill=blue!8, text width=3.2cm,
                  minimum height=0.9cm, align=center, rounded corners=2pt,
                  font=\small},
    broker/.style={rectangle, draw, fill=orange!15, text width=2.5cm,
                   minimum height=0.8cm, align=center, rounded corners=2pt,
                   font=\small},
    data/.style={rectangle, draw, fill=green!10, text width=3.2cm,
                 minimum height=0.8cm, align=center, rounded corners=2pt,
                 font=\small},
    output/.style={rectangle, draw, fill=red!10, text width=3.2cm,
                   minimum height=0.8cm, align=center, rounded corners=2pt,
                   font=\small},
    arrow/.style={-{Stealth[length=6pt]}, thick},
]

% Brokers
\node[broker] (antares) {ANTARES\\(ZTF + Rubin)};
\node[broker, right=of antares] (alerce) {ALeRCE\\(ZTF)};
\node[broker, right=of alerce] (lsst) {ALeRCE-LSST\\(Rubin)};

% Merge
\node[block, below=1.2cm of alerce] (merge) {Cross-match \&\\Deduplicate};

% Extinction
\node[block, below=of merge] (extinct) {Galactic Extinction\\(IRSA SFD)};

% Variable screening
\node[block, below=of extinct] (varscreen) {Variable Star\\Screening};

% Classification filter
\node[block, below=of varscreen] (classify) {$P(\mathrm{Ia})$ Filter\\+ Agreement};

% NED redshifts
\node[block, below=of classify] (ned) {NED Redshift\\Lookup};

% Peak fitting
\node[block, below=of ned] (peakfit) {Peak Fitting\\(Parabola + Villar)};

% Merit
\node[block, below=of peakfit] (merit) {Merit Function\\$\mathcal{M}(\Delta t, m_{\mathrm{corr}})$};

% Cache
\node[data, right=2.5cm of peakfit] (cache) {SQLite Cache\\(\texttt{peak\_fit\_targets})};

% External data
\node[data, left=2.0cm of extinct] (irsa) {IRSA Dust\\Service};
\node[data, left=2.0cm of ned] (neddb) {NED\\Database};

% Magellan
\node[output, below=of merit] (magellan) {Magellan Observing\\Catalog};

% Arrows
\draw[arrow] (antares.south) -- ++(0,-0.4) -| (merge.north);
\draw[arrow] (alerce.south) -- (merge.north);
\draw[arrow] (lsst.south) -- ++(0,-0.4) -| (merge.north);
\draw[arrow] (merge) -- (extinct);
\draw[arrow] (extinct) -- (varscreen);
\draw[arrow] (varscreen) -- (classify);
\draw[arrow] (classify) -- (ned);
\draw[arrow] (ned) -- (peakfit);
\draw[arrow] (peakfit) -- (merit);
\draw[arrow] (peakfit) -- (cache);
\draw[arrow] (cache.south) -- ++(0,-0.4) -| (magellan.east);
\draw[arrow] (merit) -- (magellan);
\draw[arrow] (irsa) -- (extinct);
\draw[arrow] (neddb) -- (ned);

\end{tikzpicture}
\caption{Pipeline data flow.  Three alert brokers feed into a
cross-matching and deduplication stage.  Galactic extinction corrections
are applied using IRSA SFD dust maps, followed by variable star screening,
classification filtering, NED spectroscopic redshift lookups, light curve
peak fitting (inverted parabola and Villar SPM), and merit scoring using
extinction-corrected magnitudes.  Results are cached in SQLite for
consumption by the Magellan planning notebook.}
\label{fig:pipeline}
\end{figure}

\subsection{Code Organization}

The codebase is organized as follows:

\begin{center}
\small
\begin{tabular}{ll}
\toprule
Module & Purpose \\
\midrule
\texttt{run\_tonight.py} & Nightly CLI: all brokers $\to$ Magellan plan \\
\texttt{supernova\_monitor.py} & Multi-broker orchestrator class \\
\texttt{broker\_clients/fink\_client.py} & Fink LSST API (Rubin photometry) \\
\texttt{broker\_clients/antares\_client.py} & ANTARES broker interface \\
\texttt{broker\_clients/alerce\_client.py} & ALeRCE REST API interface (ZTF \& LSST) \\
\texttt{broker\_clients/alerce\_db\_client.py} & ALeRCE direct PostgreSQL access \\
\texttt{broker\_clients/atlas\_client.py} & ATLAS forced photometry \\
\texttt{core/alert\_aggregator.py} & Cross-match, deduplicate, merge \\
\texttt{core/variable\_screen.py} & Known variable star screening \\
\texttt{core/ddf\_fields.py} & DDF field definitions \\
\texttt{core/peak\_fitting.py} & Peak fitting (parabola + Villar SPM) \\
\texttt{core/magellan\_planning.py} & Magellan target planning \& catalog generation \\
\texttt{cache/alert\_cache.py} & SQLite caching layer \\
\texttt{host\_galaxy/morphology\_filter.py} & Galaxy morphology filtering \\
\texttt{utils/extinction.py} & Galactic extinction corrections (IRSA SFD) \\
\texttt{utils/ned\_query.py} & NED spectroscopic redshift lookups \\
\texttt{utils/plotting.py} & Light curve visualization \\
\texttt{utils/coordinates.py} & Coordinate utilities \\
\texttt{utils/catalog\_query.py} & Catalog cross-matching utilities \\
\bottomrule
\end{tabular}
\end{center}


%% =========================================================================
\section{Data Sources}
\label{sec:data_sources}
%% =========================================================================

The pipeline queries four independent alert broker sources, each providing
different coverage, classifiers, and data products.  Candidates from all
brokers are merged, deduplicated by coordinate matching (1~arcsec tolerance),
and screened against known variable star catalogs before photometry retrieval
and light curve fitting.

\subsection{ANTARES}

ANTARES\footnote{\url{https://antares.noirlab.edu/}} ingests both ZTF
and (limited) Rubin alert streams.  The pipeline performs cone searches
of each DDF field (1.75\degr\ radius) and filters for loci with
supernova-related tags (\texttt{superphot\_plus\_classified},
\texttt{SN\_candies}, \texttt{high\_amplitude\_transient\_candidate}).

ANTARES does not provide a trained SN~Ia classifier.  Instead, the
pipeline computes a heuristic proxy $P(\mathrm{Ia})$ based on the
star/galaxy score (\texttt{ztf\_sgscore1}) and light curve duration.
This proxy is \textbf{capped at 0.50} to prevent ANTARES-only
detections from dominating the ranking:
\begin{equation}
P(\mathrm{Ia})_{\mathrm{ANTARES}} = \min\!\left(
0.50,\; f(\texttt{sgscore}, \texttt{duration})\right).
\end{equation}

\subsection{ALeRCE (ZTF)}

ALeRCE\footnote{\url{https://alerce.online/}} processes ZTF alerts
through a random-forest stamp classifier
(\texttt{lc\_classifier\_BHRF\_forced\_phot}, version~2.1.0) trained on
labeled ZTF data.  It outputs per-object probabilities for SN~Ia,
SN~II, SN~Ib/c, and SLSN classes.  Coverage is limited to the ZTF
footprint (Dec~$\gtrsim -30\degr$), so the three southern DDFs
(ELAIS-S1, EDFS\_a, EDFS\_b) are not covered.

\subsection{ALeRCE-LSST}

ALeRCE also processes Rubin/LSST alerts through a separate stamp
classifier (\texttt{stamp\_classifier\_rubin}) that classifies objects
into five categories: SN, AGN, variable star, asteroid, and bogus.
Crucially, this classifier \textbf{does not sub-classify SN types}: a
high $P(\mathrm{SN})$ indicates a likely supernova but does not
distinguish Type~Ia from Type~II or Ib/c.

\paragraph{Known limitation.} The ALeRCE LSST API does not correctly
handle cone-search queries (returns empty results).  The pipeline works
around this by querying globally for all SN-classified objects and
filtering locally by DDF coordinates using \texttt{astropy}.

ALeRCE-LSST provides the richest Rubin photometry: multi-epoch
$ugrizy$ light curves with both alert-triggered detections and forced
photometry at the object position.

\subsection{Fink}

Fink\footnote{\url{https://fink-portal.org/}} processes Rubin/LSST
alerts through multiple machine-learning classifiers including
\texttt{snnSnVsOthers} (SN vs.\ non-SN) and \texttt{earlySNIa} (early
Type~Ia identification).  The pipeline queries the Fink LSST
API\footnote{\url{https://api.lsst.fink-portal.org/}} for candidates
tagged as \texttt{sn\_near\_galaxy\_candidate} and
\texttt{extragalactic\_new\_candidate}, filtering by Fink's SN
classification score.

Fink serves as the primary source of Rubin prompt photometry: for each
candidate, the pipeline retrieves the full multi-band ($grizy$) light
curve including DiaSource detections and forced photometry at every
visit.  Flux is provided directly in nanoJanskys from Rubin difference
imaging.  No authentication is required.

\subsection{ALeRCE Direct Database Access}
\label{sec:alerce_db}

In addition to the REST API, the pipeline can query ALeRCE's ZTF
database directly via PostgreSQL for bulk data retrieval.  Read-only
credentials are fetched at runtime from the ALeRCE
\texttt{usecases} repository on
GitHub.\footnote{\url{https://github.com/alercebroker/usecases}}

The \texttt{AlerceDBClient} class
(\texttt{broker\_clients/alerce\_db\_client.py}) provides methods for:
\begin{itemize}
\item Bulk SN candidate queries with classifier probability thresholds,
\item SPM (Villar supernova parametric model) feature retrieval,
\item Per-object classification probabilities (pivoted by class),
\item Pan-STARRS host galaxy photometry (\texttt{sgscore1}, \texttt{sgmag1}, colors),
\item Light curve detections (\texttt{mjd}, \texttt{magpsf}, \texttt{sigmapsf}).
\end{itemize}

All queries operate on ALeRCE's \texttt{object}, \texttt{probability},
\texttt{feature}, \texttt{detection}, and \texttt{ps1\_ztf} tables.  The
direct database path is significantly faster than the REST API for batch
queries returning hundreds to thousands of objects.  If the database is
unreachable (e.g., no \texttt{psycopg2} installation or network
restrictions), the pipeline falls back to the REST API transparently.

\begin{table}[h]
\centering
\caption{Summary of alert broker properties.}
\label{tab:brokers}
\begin{tabular}{lcccl}
\toprule
Broker & Survey & Dec limit & Classifier & $P(\mathrm{Ia})$ range \\
\midrule
ANTARES       & ZTF + Rubin & $\gtrsim -30\degr$ (ZTF) & Heuristic proxy & 0.05--0.50 \\
ALeRCE (ZTF)  & ZTF         & $\gtrsim -30\degr$       & RF stamp classifier & 0.0--1.0 \\
ALeRCE-LSST   & Rubin       & All-sky                  & Rubin stamp classifier & 0.0--1.0 \\
\bottomrule
\end{tabular}
\end{table}


%% =========================================================================
\section{Alert Aggregation}
\label{sec:aggregation}
%% =========================================================================

The \texttt{AlertAggregator} class (\texttt{core/alert\_aggregator.py})
merges detections from all three brokers into a single, deduplicated
candidate list.

\subsection{Cross-Matching and Deduplication}

Objects from different brokers are matched using two criteria:
\begin{enumerate}
\item \textbf{Coordinate proximity:} Two detections are considered the
same astrophysical source if their angular separation is
$\leq 1\arcsec$ (converted to degrees: $1/3600 \approx
0.000278\degr$).
\item \textbf{Shared identifiers:} Objects sharing a ZTF
\texttt{object\_id} across brokers are merged regardless of coordinate
offset.
\end{enumerate}

For each unique source, the pipeline creates a merged record retaining
broker-specific identifiers (\texttt{object\_id\_ANTARES},
\texttt{object\_id\_ALeRCE}, \texttt{object\_id\_ALeRCE-LSST}),
classification probabilities, and quality diagnostics.

\subsection{Probability Normalization}

Each broker's classification probability is normalized to a common
$P(\mathrm{Ia})$ scale:
\begin{itemize}
\item \textbf{ALeRCE (ZTF):} Direct ML probability used as-is.
\item \textbf{ALeRCE-LSST:} $P(\mathrm{SN})$ used as a proxy for
$P(\mathrm{Ia})$, acknowledging the lack of subtype discrimination.
\item \textbf{ANTARES:} Heuristic proxy capped at 0.50 (see
\S\ref{sec:data_sources}).
\end{itemize}

\subsection{Agreement Scoring}

For objects detected by multiple brokers, the pipeline computes
aggregate statistics across all available $P(\mathrm{Ia})$ values:
\begin{align}
\bar{P} &= \frac{1}{N}\sum_{i=1}^{N} P_i(\mathrm{Ia}), \\
\sigma_P &= \mathrm{std}(P_1, \ldots, P_N).
\end{align}

An agreement score is assigned based on $\sigma_P$:
\begin{equation}
\mathrm{Agreement} =
\begin{cases}
1.0 & \text{if } \sigma_P < 0.1 \quad \text{(high)}, \\
0.5 & \text{if } 0.1 \leq \sigma_P < 0.25 \quad \text{(moderate)}, \\
0.0 & \text{if } \sigma_P \geq 0.25 \quad \text{(low)}.
\end{cases}
\end{equation}

\subsection{Variable Star Screening}

The \texttt{VariableScreener} class cross-matches candidates against
catalogs of 13{,}749 known variable stars compiled from Gaia~DR3 and
the AAVSO Variable Star Index (VSX), distributed across the six Rubin
DDFs.  The matching radius is 2\arcsec.  Matched candidates are flagged
with \texttt{known\_variable = True} and excluded from the SN~Ia
candidate list.


%% =========================================================================
\section{Galactic Extinction Corrections}
\label{sec:extinction}
%% =========================================================================

Galactic dust along the line of sight reddens and dims extragalactic
sources.  Correcting for this extinction is essential for accurate peak
magnitude estimates and, consequently, for meaningful merit scores and
absolute magnitude calculations.

\subsection{Dust Map Queries}

The pipeline queries the IRSA Galactic Dust Reddening and Extinction
service\footnote{\url{https://irsa.ipac.caltech.edu/applications/DUST/}}
using the \texttt{astroquery.ipac.irsa.irsa\_dust.IrsaDust} interface.
For each candidate, the \citet{Schlafly2011} recalibration of the
\citet{Schlegel1998} (SFD) dust map provides the total Galactic
extinction $A_\lambda$ in each photometric band:
\begin{equation}
A_\lambda(\ell, b) = R_\lambda \times E(B-V)_{\mathrm{SFD}},
\end{equation}
where $R_\lambda$ is the ratio of total-to-selective extinction for
band~$\lambda$ and $E(B-V)_{\mathrm{SFD}}$ is the reddening from the
SFD map.

The service returns extinction values $A_{\mathrm{SFD}}$ for SDSS
$ugriz$ filters.  The pipeline extracts $A_g$, $A_r$, $A_i$, $A_z$ for
each candidate based on its individual sky coordinates (not a per-field
average).

\subsection{Caching}

Extinction values are cached in the SQLite database
(\texttt{galaxy\_info} table, \texttt{extinction\_json} column) keyed by
coordinates with a tolerance of $0.1\arcmin$.  Subsequent pipeline runs
retrieve cached values without re-querying IRSA, significantly reducing
runtime for repeated analyses of the same fields.

\subsection{Extinction-Corrected Magnitudes}

The corrected apparent magnitude for band~$\lambda$ is:
\begin{equation}
m_{\lambda,\mathrm{corr}} = m_\lambda - A_\lambda,
\end{equation}
where $m_\lambda$ is the observed magnitude and $A_\lambda$ is the
Galactic extinction in that band.  The extinction-corrected peak
magnitudes ($\texttt{peak\_mag\_corrected}$) propagate through the merit
function and Magellan catalog generation.

\subsection{Typical Extinction Values}

The DDF fields span a range of Galactic latitudes:

\begin{center}
\begin{tabular}{lcl}
\toprule
Field & Galactic $|b|$ & Typical $A_g$ \\
\midrule
COSMOS & $42\degr$ & $\sim 0.07$~mag \\
M49 & $70\degr$ & $\sim 0.08$~mag \\
XMM-LSS & $-59\degr$ & $\sim 0.10$~mag \\
ECDFS & $-55\degr$ & $\sim 0.03$~mag \\
ELAIS-S1 & $-73\degr$ & $\sim 0.04$~mag \\
\bottomrule
\end{tabular}
\end{center}

While individual corrections are modest for most DDFs, they vary
across each field and are important for precision photometry at the
$\lesssim 0.1$~mag level relevant to SN~Ia standardization.


%% =========================================================================
\section{NED Spectroscopic Redshifts}
\label{sec:ned}
%% =========================================================================

For each SN candidate, the pipeline queries the NASA/IPAC
Extragalactic Database
(NED)\footnote{\url{https://ned.ipac.caltech.edu/}} for spectroscopic
redshifts of potential host galaxies.  Spectroscopic redshifts, when
available, provide accurate distance estimates and enable absolute
magnitude calculations.

\subsection{Query Strategy}

The \texttt{utils/ned\_query.py} module uses
\texttt{astroquery.ipac.ned.Ned.query\_region()} to search for
cataloged objects within a radius of 18\arcsec\ of each SN candidate
position.  This search radius is chosen to encompass typical SN--host
offsets while limiting contamination from unrelated sources.

From the returned objects, the pipeline selects the nearest source with
a valid spectroscopic redshift.  The result includes:
\begin{itemize}
\item \texttt{ned\_redshift}: spectroscopic redshift $z$,
\item \texttt{ned\_name}: NED designation of the host galaxy,
\item \texttt{ned\_sep\_arcsec}: angular separation from the SN position,
\item \texttt{ned\_distmod}: distance modulus $\mu$.
\end{itemize}

\subsection{Distance Modulus}

The distance modulus is computed using:
\begin{equation}
\mu = 5 \log_{10}\!\left(\frac{d_L}{10\,\mathrm{pc}}\right),
\end{equation}
where $d_L$ is the luminosity distance derived from the redshift using
WMAP7 cosmology ($H_0 = 70.4$~km\,s$^{-1}$\,Mpc$^{-1}$,
$\Omega_m = 0.272$).

\subsection{Absolute Magnitude Estimation}

When both extinction-corrected peak magnitude and NED distance modulus
are available, the pipeline estimates the absolute magnitude:
\begin{equation}
M = m_{\mathrm{corr}} - \mu.
\end{equation}
For Type~Ia supernovae, the expected range is $M \approx -19.3 \pm
0.5$~mag.  Candidates with anomalous absolute magnitudes (e.g.,
$M < -21$ or $M > -16$) may warrant additional scrutiny or may indicate
misclassification.

\subsection{Caching and Rate Limiting}

NED query results are cached in the \texttt{galaxy\_info} table (columns
\texttt{redshift}, \texttt{ned\_name}, \texttt{ned\_separation\_arcsec}).
Batch queries iterate over unique coordinates with error handling for
timeouts and empty results.  NED's query service can be slow for large
batches, so caching is important for interactive use.


%% =========================================================================
\section{Rubin Difference-Imaging Photometry}
\label{sec:photometry}
%% =========================================================================

ALeRCE-LSST photometry is derived from Rubin \emph{difference imaging}:
each science image is subtracted from a co-added reference template, and
point-spread function (PSF) photometry is performed on the residual.

\subsection{Flux Representation}

The difference-image flux (\texttt{psfFlux}) is reported in
\textbf{nanoJanskys} (nJy) and represents the flux difference
$f_{\mathrm{sci}} - f_{\mathrm{tmpl}}$:
\begin{itemize}
\item $f > 0$: source is brighter than template (transient brightening).
\item $f < 0$: source is fainter than template (fading below baseline).
\item $f = 0$: source matches the template exactly.
\end{itemize}

The science-image flux (\texttt{scienceFlux}) is also available for
alert-triggered detections (but not for forced photometry) and is always
positive for real sources.

\subsection{Magnitude Conversion}

For positive flux values, the AB magnitude is:
\begin{equation}
\label{eq:mag_conversion}
m = -2.5 \log_{10}(f_{\mathrm{nJy}}) + 31.4,
\end{equation}
with uncertainty propagated as:
\begin{equation}
\label{eq:mag_err}
\sigma_m = \frac{2.5}{\ln 10} \cdot \frac{\sigma_f}{f}.
\end{equation}

\paragraph{Negative flux limitation.} Points with $f \leq 0$ cannot
be converted to magnitudes and are dropped in magnitude-space
representations.  This truncates the light curve tail, which is why all
fitting in this pipeline is performed in flux space.

\subsection{Forced Photometry}

Forced photometry provides flux measurements at the object position even
when the source is not detected above threshold, giving more complete
temporal coverage than alert-triggered detections alone.  However, forced
photometry only provides \texttt{psfFlux} (difference flux); the
\texttt{scienceFlux} is not available.


%% =========================================================================
\section{Light Curve Peak Fitting}
\label{sec:peak_fitting}
%% =========================================================================

The \texttt{core/peak\_fitting.py} module estimates the time of peak
brightness and peak apparent magnitude for each candidate.  Two methods
are available.

\subsection{Inverted Parabola Model}

Near maximum light, the flux light curve of a SN~Ia is approximately
parabolic.  We fit an inverted parabola in flux space:
\begin{equation}
\label{eq:parabola}
f(t) = f_{\mathrm{peak}} - a\,(t - t_0)^2,
\end{equation}
where $f_{\mathrm{peak}}$ is the peak flux (nJy), $t_0$ is the time of
peak (MJD), and $a > 0$ is the curvature parameter.

The fit is performed independently per band using
\texttt{scipy.optimize.curve\_fit} with the flux uncertainties as
weights (absolute sigma).  Initial guesses are derived from the data:
$f_{\mathrm{peak}}^{(0)} = \max(f)$, $t_0^{(0)} = t(\max f)$,
$a^{(0)} = 1.0$.

\paragraph{Working in flux space.} Fitting in flux space rather than
magnitude space has three advantages: (1)~negative difference-imaging
flux values are handled naturally; (2)~the parabola is a more physical
model near peak in linear flux than in logarithmic magnitudes; (3)~there
are no singularities from the $\log_{10}$ transform.

\subsection{Goodness of Fit}

The reduced chi-squared is:
\begin{equation}
\chi^2_\nu = \frac{1}{\nu} \sum_{i=1}^{N}
\left(\frac{f_i - f(t_i)}{\sigma_{f,i}}\right)^2,
\end{equation}
where $\nu = \max(N-3, 1)$ is the number of degrees of freedom (three
free parameters: $f_{\mathrm{peak}}$, $t_0$, $a$).

\subsection{Fit Status Flags}

Each per-band fit is assigned a status flag:

\begin{center}
\begin{tabular}{ll}
\toprule
Status & Criterion \\
\midrule
\texttt{ok} & $N \geq 4$ and $\chi^2_\nu \leq 10$ \\
\texttt{underdetermined} & $N = 3$ (exactly constrained) \\
\texttt{poor\_fit} & $N \geq 4$ and $\chi^2_\nu > 10$ \\
\texttt{insufficient\_data} & $N < 3$ \\
\texttt{fit\_failed} & Optimizer did not converge \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Band Priority}

For the ``headline'' peak estimate reported per candidate, the best band
is selected from the priority list:
\begin{equation}
r > g > i > z > y > u.
\end{equation}
The $r$-band is prioritized because it is closest to rest-frame $B$-band
for SN~Ia at the typical DDF redshift range ($z \sim 0.1$--$0.4$).  The
first band with a successful fit (status \texttt{ok} or
\texttt{underdetermined}) is selected.

\subsection{Multi-Band Villar Supernova Parametric Model}
\label{sec:villar}

The primary peak-fitting method is a simultaneous multi-band fit of
the Villar supernova parametric model \citep{Villar2019} directly to
the combined Rubin\,+\,ZTF\,+\,ATLAS light curve data.  This fit
uses a shared explosion epoch $t_0$ across all bands while allowing
per-band amplitude and timescale parameters, providing much more
robust peak estimates than independent per-band fits.

\subsubsection{Model Definition}

The Villar model describes the flux as a function of time with a smooth
sigmoid rise and exponential decline:
\begin{equation}
\label{eq:villar}
F(t) =
\begin{cases}
\displaystyle
\frac{A\!\left[1 - \beta\,\frac{t - t_0}{t_0 + \gamma - t_0}\right]}
{1 + \exp\!\left(-\frac{t - t_0}{\tau_{\mathrm{rise}}}\right)},
& t < t_0 + \gamma, \\[12pt]
\displaystyle
\frac{A(1-\beta)\,\exp\!\left(-\frac{t - t_0 - \gamma}{\tau_{\mathrm{fall}}}\right)}
{1 + \exp\!\left(-\frac{t - t_0}{\tau_{\mathrm{rise}}}\right)},
& t \geq t_0 + \gamma,
\end{cases}
\end{equation}
where $t_0$ is the explosion epoch (MJD), $A$ is the peak amplitude
in nanoJanskys, $\beta$ controls the pre-maximum slope,
$\gamma$ sets the rise-to-decline transition time,
$\tau_{\mathrm{rise}}$ and $\tau_{\mathrm{fall}}$ are the rise and
fall timescales in days.

\subsubsection{Multi-Band Fitting Strategy}

The key innovation is sharing $t_0$ across all photometric bands.
For $N_b$ bands, the parameter vector is:
\begin{equation}
\vec{p} = \bigl[t_0,\; A_1, \beta_1, \gamma_1, \tau_{\mathrm{rise},1},
\tau_{\mathrm{fall},1},\; A_2, \beta_2, \ldots\bigr],
\end{equation}
with $1 + 5N_b$ total free parameters.  The fit minimizes
\begin{equation}
\chi^2 = \sum_{b=1}^{N_b} \sum_{i=1}^{n_b}
\left(\frac{F_{b,i}^{\mathrm{obs}} - F_b(t_i;\,\vec{p})}
{\sigma_{b,i}}\right)^2
\end{equation}
using \texttt{scipy.optimize.least\_squares} with the Trust Region
Reflective algorithm and Huber-like \texttt{soft\_l1} robust loss
function, which down-weights outliers.  Bounds constrain all
parameters to physically reasonable ranges (e.g., $A > 0$,
$0 < \beta < 1$, $\gamma > 1$~day).

This multi-band approach constrains the explosion epoch using data
from every available band simultaneously, which is particularly
valuable for the Rubin DDFs where $g$, $r$, $i$, $z$ photometry
is available plus supplementary ZTF $g$/$r$ and ATLAS cyan/orange
data when the target is in the northern sky.

\subsubsection{Light Curve Cleaning}

Before fitting, the \texttt{clean\_light\_curve()} function processes
the combined photometry:
\begin{enumerate}
\item \textbf{Deduplication:} When a DiaSource detection and a
forced-photometry measurement exist at the same (MJD, band), the
detection is kept (it has better-characterized uncertainties).
\item \textbf{Sigma clipping:} Per (night, band) group, points
deviating more than $4\sigma$ from the inverse-variance weighted
mean flux are rejected.  This removes catastrophic outliers from
difference-imaging artifacts.
\end{enumerate}

\subsubsection{Photometric Quality Cuts}

Before fitting, each candidate must satisfy three criteria designed to
reject single-epoch events, single-band artifacts, and poorly-sampled
objects that cannot be reliably typed:
\begin{enumerate}
\item At least 5 photometric data points with $\mathrm{SNR} > 5$.
\item Detections (among high-SNR points) in at least 2 photometric
  bands.
\item A time baseline of at least 2~days between first and last
  observation.
\end{enumerate}
After fitting, the multi-band Villar SPM model must converge with
data from $\geq 2$ bands.  If the Villar fit fails, the inverted
parabola is accepted only when it too succeeds in $\geq 2$ bands.

\subsubsection{Numerical Peak Finding}

The time of peak brightness is determined by numerical maximization of
the flux function rather than the analytic approximation $t_{\mathrm{peak}}
\approx t_0 + \gamma$.  The analytic formula can be
inaccurate by 7--21~days because the sigmoid rise term shifts the true
flux maximum away from the transition time.

The pipeline evaluates $F(t)$ on a fine grid of 2000 points spanning
$[t_0,\; t_0 + \gamma + 3\tau_{\mathrm{fall}}]$
and identifies the grid point with maximum flux:
\begin{equation}
t_{\mathrm{peak}} = \arg\max_t F(t).
\end{equation}
The corresponding peak magnitude is:
\begin{equation}
m_{\mathrm{peak}} = -2.5\,\log_{10}(F_{\mathrm{peak}}/\mathrm{nJy}) + 31.4\,\mathrm{mag},
\end{equation}
where $31.4 = -2.5\log_{10}(3.631\times10^{-6}) \approx 8.9 + 22.5$ is
the AB zero-point for nanoJansky flux units.

\subsubsection{Multi-Survey Photometry Combination}

The nightly pipeline (\texttt{run\_tonight.py}) combines photometry
from up to three surveys before fitting:
\begin{itemize}
\item \textbf{Rubin/LSST} (via Fink): difference-imaging photometry
  in $g$, $r$, $i$, $z$ bands, including both DiaSource detections
  and forced photometry.  Flux in nanoJanskys (nJy).
\item \textbf{ZTF} (via ALeRCE): detections in $g$, $r$ bands.
  Magnitudes converted to nJy via
  $F_{\mathrm{nJy}} = 10^{(31.4 - m)/2.5}$.
\item \textbf{ATLAS}: forced photometry in cyan ($c$) and orange ($o$)
  bands, queried only for candidates brighter than 20th magnitude
  (the approximate ATLAS detection limit).  Targets are submitted
  in batch via the \texttt{radeclist} API endpoint for efficiency.
  Flux in microJanskys converted via
  $F_{\mathrm{nJy}} = F_{\mu\mathrm{Jy}} \times 1000$.
\end{itemize}
All photometry is standardized to a common DataFrame with columns
\texttt{mjd}, \texttt{flux}, \texttt{flux\_err}, \texttt{band},
\texttt{survey} before the combined fit.

\subsubsection{Integration with Parabola Fits}

Both the multi-band Villar fit and per-band parabola fits are run on
every candidate.  The Villar result is preferred when it converges
(fit status = \texttt{ok}), with the parabola as fallback.  The
summary table includes peak magnitude, peak MJD, delta-$t$ (days
since peak), and the fit method used for each candidate.

\subsection{Optional: SALT2 Template Fitting}

If the \texttt{sncosmo} package is installed, a multi-band SALT2
template fit is also available.  The SALT2 model parameterizes SN~Ia
light curves as:
\begin{equation}
f(t, \lambda) = x_0 \left[ M_0(t, \lambda) + x_1\, M_1(t, \lambda)
\right] \times \exp\!\left[ c\, \mathrm{CL}(\lambda) \right],
\end{equation}
where $x_0$ is the amplitude, $x_1$ is the stretch parameter
(light-curve width), $c$ is the color parameter, $M_0$ and $M_1$ are
the template spectral surfaces, and $\mathrm{CL}(\lambda)$ is the color
law.  The fit simultaneously uses all LSST bands ($g$, $r$, $i$, $z$,
$y$, $u$ mapped to \texttt{sncosmo} band names \texttt{lsstg},
\texttt{lsstr}, etc.) and requires $\geq 5$ data points across $\geq 2$
bands.

The SALT2 fit returns $t_0$ (time of $B$-band maximum), $x_1$
(stretch), $c$ (color), the peak rest-frame $B$-band magnitude, and
the redshift $z$ (which can float or be fixed if a spectroscopic
redshift is known).


%% =========================================================================
\section{Merit Function for Spectroscopic Follow-Up}
\label{sec:merit}
%% =========================================================================

Given the fitted peak time $t_0$ and extinction-corrected peak magnitude
$m_{\mathrm{corr}}$, we define a merit function that encodes the
scientific value of spectroscopically observing a candidate on a given
night:
\begin{equation}
\label{eq:merit}
\boxed{\mathcal{M} = W_t(\Delta t) \times W_m(m),}
\end{equation}
where $\Delta t = t_{\mathrm{now}} - t_0$ is the time since peak (days,
positive means post-peak), and $m$ is the extinction-corrected peak
apparent magnitude $m_{\mathrm{corr}}$ (falling back to the uncorrected
$m$ if extinction data are unavailable).

\subsection{Time Weight}

The time weight is a Gaussian centered on peak:
\begin{equation}
\label{eq:wt}
W_t = \exp\!\left(-\frac{\Delta t^2}{2\tau^2}\right),
\end{equation}
with $\tau = 10$~days by default.  This gives:
\begin{center}
\begin{tabular}{rc}
\toprule
$|\Delta t|$ (days) & $W_t$ \\
\midrule
0 & 1.000 \\
5 & 0.882 \\
10 & 0.607 \\
15 & 0.325 \\
20 & 0.135 \\
30 & 0.011 \\
\bottomrule
\end{tabular}
\end{center}

The Gaussian is symmetric: pre-peak observations (rising light curve)
receive the same weight as post-peak.  This is appropriate because
pre-peak spectra are highly valuable for constraining explosion physics,
while post-peak spectra within $\sim$2~weeks are standard for
classification.

\subsection{Magnitude Weight}

The magnitude weight is a Gaussian in apparent magnitude space:
\begin{equation}
\label{eq:wm}
W_m = \exp\!\left(-\frac{(m - m_{\mathrm{opt}})^2}{\sigma_m^2}\right),
\end{equation}
with default parameters:
\begin{itemize}
\item $m_{\mathrm{opt}} = 20.5$ --- the optimal magnitude for Magellan
6.5\,m spectroscopy.  At this brightness, a SN~Ia at $z \sim 0.1$--0.15
yields a classification-quality spectrum in $\sim$20--40 minutes with
LDSS3C or IMACS.
\item $\sigma_m = \frac{m_{\mathrm{faint}} - m_{\mathrm{bright}}}{4} =
\frac{23.0 - 18.0}{4} = 1.25$~mag --- chosen so that the weight spans
the useful magnitude range ($\sim$18--23~mag) with reasonable falloff.
\end{itemize}

\subsection{Combined Merit Surface}

The product $\mathcal{M} = W_t \times W_m$ defines a two-dimensional
surface over $(\Delta t, m)$ space.  The maximum merit
($\mathcal{M} = 1$) occurs at $\Delta t = 0$, $m = m_{\mathrm{opt}}$.
Objects far from peak, or too bright (nearby, less cosmologically
interesting) or too faint (requiring prohibitively long exposures) are
deprioritized.

Candidates with $\texttt{peak\_mjd} = \mathrm{NaN}$ or
$\texttt{peak\_mag} = \mathrm{NaN}$ (due to insufficient photometry or
fit failure) receive $\mathcal{M} = \mathrm{NaN}$ and are placed at
the bottom of the target list.

All merit function parameters ($\tau$, $m_{\mathrm{opt}}$,
$m_{\mathrm{bright}}$, $m_{\mathrm{faint}}$) are configurable to
accommodate different instruments and science goals.


%% =========================================================================
\section{Magellan Observing Plan Generation}
\label{sec:magellan}
%% =========================================================================

The final stage of the pipeline generates a target catalog in the native
Magellan Telescope Control System (TCS) format for spectroscopic
follow-up.

\subsection{Observability Constraints}

For a specified UT observing date, the pipeline computes target
observability from Las Campanas Observatory:
\begin{center}
\begin{tabular}{ll}
Latitude:  & $-29.0146\degr$ \\
Longitude: & $-70.6926\degr$ \\
Altitude:  & 2516\,m \\
\end{tabular}
\end{center}

Using \texttt{astropy.coordinates.AltAz} and
\texttt{astropy.coordinates.get\_sun}, the pipeline:
\begin{enumerate}
\item Computes nautical twilight boundaries (Sun altitude $< -12\degr$).
\item Creates a time grid at 15-minute intervals spanning the dark hours.
\item For each target, computes altitude at each grid point.
\item Retains targets that spend at least a configurable minimum time
(default: 1~hour) above the airmass limit (default: airmass $< 2.0$,
corresponding to altitude $> 30\degr$).
\end{enumerate}

\subsection{Target Prioritization}

Observable targets are sorted by Right Ascension (ascending) for
efficient telescope slewing.  Within the catalog, comment lines encode
the merit score, peak magnitude, time from peak, DDF field, and fit
status for each target, enabling the observer to make informed decisions
about which targets to prioritize during the night.

\subsection{Catalog Format}

The Magellan TCS requires a 16-field whitespace-separated format:

\begin{center}
\small
\begin{tabular}{rlp{7cm}}
\toprule
Field & Name & Description \\
\midrule
1  & Reference number & Sequential integer \\
2  & Object name & Alphanumeric $+$ \texttt{+-.\textunderscore} only \\
3  & RA & \texttt{hh:mm:ss.s} (J2000) \\
4  & Dec & \texttt{$\pm$dd:mm:ss} (J2000) \\
5  & Equinox & \texttt{2000.0} \\
6  & PM$_\alpha$ & 0.000 (no proper motion) \\
7  & PM$_\delta$ & 0.000 \\
8  & Rotator angle & Instrument-dependent (degrees) \\
9  & Rotator mode & \texttt{GRV}, \texttt{EQU}, \texttt{HRZ}, or \texttt{OFF} \\
10--12 & Guide probe 1 & RA, Dec, equinox (zeros if unused) \\
13--15 & Guide probe 2 & RA, Dec, equinox (zeros if unused) \\
16 & Epoch & 0.000 (no proper motion) \\
\bottomrule
\end{tabular}
\end{center}

Instrument presets (LDSS3C, IMACS, MagE, FIRE) provide default rotator
angles and modes.  Guide probe fields are set to zero (no preselected
guide stars).


%% =========================================================================
\section{Data Schema}
\label{sec:schema}
%% =========================================================================

The pipeline uses a SQLite database (\texttt{cache/data/alerts\_cache.db})
for caching intermediate results and bridging between the discovery
notebook and the Magellan planning notebook.

\subsection{Existing Tables}

\paragraph{\texttt{alerts}} Caches raw broker query results.  Columns:
\texttt{object\_id} (TEXT), \texttt{broker} (TEXT), \texttt{ra} (REAL),
\texttt{dec} (REAL), \texttt{discovery\_date} (TEXT),
\texttt{classification} (JSON), \texttt{metadata} (JSON),
\texttt{cached\_at} (TIMESTAMP).  Unique on
\texttt{(object\_id, broker)}.

\paragraph{\texttt{galaxy\_info}} Caches galaxy morphology, photometry,
extinction, and redshift data from catalog queries.  Columns:
\texttt{ra}, \texttt{dec}, \texttt{morphology}, \texttt{catalog} (JSON),
\texttt{redshift}, \texttt{magnitude\_g/r/i/z},
\texttt{extinction\_json} (JSON dict of $\{$band: $A_\lambda\}$ from
IRSA SFD), \texttt{ned\_name} (NED host designation),
\texttt{ned\_separation\_arcsec} (angular offset from SN position).
Unique on \texttt{(ra, dec)}.

\paragraph{\texttt{merged\_alerts}} Caches deduplicated/merged alerts.
Columns: \texttt{unique\_id}, \texttt{ra}, \texttt{dec},
\texttt{brokers\_detected}, per-broker classification JSON,
\texttt{last\_updated}.

\subsection{New Table: \texttt{peak\_fit\_targets}}

This table serves as the intermediate data product between the main
pipeline and the Magellan planning notebook:

\begin{center}
\small
\begin{tabular}{lll}
\toprule
Column & Type & Description \\
\midrule
\texttt{object\_id}              & TEXT  & Primary identifier \\
\texttt{object\_id\_alerce\_lsst}& TEXT  & ALeRCE-LSST ID \\
\texttt{object\_id\_antares}     & TEXT  & ANTARES ID \\
\texttt{ddf\_field}              & TEXT  & DDF field name \\
\texttt{ra}, \texttt{dec}        & REAL  & J2000 coordinates (deg) \\
\texttt{mean\_ia\_prob}          & REAL  & Mean $P(\mathrm{Ia})$ \\
\texttt{brokers\_detected}       & TEXT  & Broker names \\
\texttt{A\_g}, \texttt{A\_r}, \texttt{A\_i} & REAL & Galactic extinction (mag) \\
\texttt{ned\_redshift}           & REAL  & NED spectroscopic redshift \\
\texttt{ned\_distmod}            & REAL  & Distance modulus from NED $z$ \\
\texttt{peak\_mjd}               & REAL  & Fitted peak time (MJD) \\
\texttt{peak\_mag}               & REAL  & Fitted peak magnitude (AB) \\
\texttt{peak\_mag\_corrected}    & REAL  & Extinction-corrected peak mag \\
\texttt{peak\_mag\_err}          & REAL  & Magnitude uncertainty \\
\texttt{peak\_band}              & TEXT  & Best-fit band \\
\texttt{peak\_fit\_status}       & TEXT  & Fit quality flag \\
\texttt{villar\_peak\_mjd}       & REAL  & Villar SPM peak time (MJD) \\
\texttt{villar\_peak\_mag}       & REAL  & Villar SPM peak magnitude \\
\texttt{mjd\_now}                & REAL  & MJD when merit computed \\
\texttt{delta\_t}                & REAL  & $t_{\mathrm{now}} - t_0$ (days) \\
\texttt{merit}                   & REAL  & Merit score (0--1) \\
\texttt{abs\_mag}                & REAL  & $m_{\mathrm{corr}} - \mu$ \\
\texttt{cached\_at}              & TIMESTAMP & Write timestamp \\
\bottomrule
\end{tabular}
\end{center}

The table uses \texttt{INSERT OR REPLACE} keyed on
\texttt{object\_id}, so re-running the main pipeline overwrites stale
entries.


%% =========================================================================
\section{Current Status and Future Work}
\label{sec:status}
%% =========================================================================

\subsection{Implemented}

\begin{itemize}
\item Multi-broker alert ingestion (ANTARES, ALeRCE-ZTF, ALeRCE-LSST)
with DDF-targeted queries and local caching.
\item ALeRCE direct PostgreSQL database access for bulk candidate
queries, SPM features, classification probabilities, and Pan-STARRS
host photometry, with transparent fallback to the REST API.
\item Cross-matching and deduplication across brokers (1\arcsec\
coordinate tolerance $+$ shared ZTF ID matching).
\item Galactic extinction corrections using IRSA SFD dust maps, with
per-object lookups and SQLite caching (see \S\ref{sec:extinction}).
\item NED spectroscopic redshift lookups for host galaxies, with
distance modulus computation and absolute magnitude estimation (see
\S\ref{sec:ned}).
\item Variable star screening against 13{,}749 known variables from
Gaia~DR3 and VSX.
\item Classification filtering with configurable $P(\mathrm{Ia})$
threshold, multi-broker agreement scoring, and minimum broker count.
\item Optional elliptical host galaxy morphology filtering via
cross-match with SDSS and Pan-STARRS galaxy catalogs (color cuts:
$g-r > 0.55$ and $r-i > 0.15$).
\item ATLAS forced photometry enrichment (cyan and orange bands) for
independent light curve validation.
\item Flux-space and magnitude-space light curve visualization with
per-band colors and survey-specific markers.
\item LSST alert stamp display (Science/Template/Difference cutouts).
\item Interactive human-in-the-loop visual classification tool
(SN/variable/bogus labels saved to CSV).
\item Inverted-parabola light curve peak fitting in flux space with
per-band fits, band-priority selection, and batch processing.
\item Villar SPM peak fitting using pre-computed ALeRCE features with
numerical flux maximization for accurate peak time determination
(see \S\ref{sec:villar}).
\item Optional SALT2 template fitting (gated by \texttt{sncosmo}
availability).
\item Two-dimensional merit function $\mathcal{M}(\Delta t, m_{\mathrm{corr}})$
using extinction-corrected peak magnitudes.
\item SQLite caching of peak-fit targets with merit scores
(\texttt{peak\_fit\_targets} table).
\item Magellan observing catalog generation with observability filtering
from Las Campanas, RA-ordered target lists, and instrument-specific
rotator presets (LDSS3C, IMACS, MagE, FIRE).
\item CSV export of candidate tables with timestamps.
\item Diagnostic plots: light curves with Villar SPM overlays and peak
markers, discovery space ($\Delta t$ vs.\ peak magnitude with merit
color-coding), extinction histograms, redshift distributions, and
Hubble diagrams.
\item Pipeline logging: every run writes a \texttt{DEBUG}-level log
file (\texttt{pipeline.log}) to the night output directory, capturing
all warnings, errors, and timing for post-run diagnostics.
\item Fink API circuit breaker: after five consecutive photometry
request failures (each timing out at 60\,s), the pipeline skips
remaining Fink queries to avoid hanging when the server is
unresponsive.
\end{itemize}

\subsection{Planned}

\begin{itemize}
\item Standalone Magellan planning notebook reading from the SQLite
cache, with airmass charts and sky-map visualizations.
\item Exposure time estimation based on target magnitude and desired
signal-to-noise ratio.
\item Photometric redshift estimation from multi-band light curve
fitting (e.g., via SALT2 with floating $z$).
\item Integration with Rubin alert stream when LSST operations begin.
\end{itemize}

\subsection{Known Limitations}

\begin{enumerate}
\item ALeRCE-LSST cone search API is broken; global query with local
filtering is used as a workaround.
\item LSST stamp classifier does not sub-classify SN types ($P(\mathrm{SN})$
only, not $P(\mathrm{Ia})$).
\item ANTARES $P(\mathrm{Ia})$ is a heuristic proxy, not an ML
classifier, capped at 0.50.
\item Negative difference-imaging flux is lost in magnitude-space
representations; all fitting is performed in flux space.
\item The 1\arcsec\ cross-match tolerance may occasionally merge
distinct sources in crowded fields.
\item LSST \texttt{scienceFlux} is only available for alert detections,
not forced photometry.
\item The inverted-parabola model is only valid near peak; the
multi-band Villar SPM fit provides a better full light curve model
but requires well-sampled multi-band data ($\geq 5$ points per band).
\item Pan-STARRS and SDSS host galaxy data are only available for DDFs
at Dec~$\gtrsim -30\degr$ (COSMOS, XMM-LSS, M49, and marginally
ECDFS).  The three southern DDFs (ELAIS-S1, EDFS\_a, EDFS\_b) lack
coverage for elliptical host classification using optical colors.
\item NED spectroscopic redshift coverage is incomplete; only
$\sim$10--30\% of SN candidates have NED-cataloged host galaxies with
measured redshifts, depending on the field depth.
\item At the default ALeRCE classifier threshold of $P > 0.5$,
superluminous supernovae (SLSN) dominate the high-probability tail
because they are more distinctive to the classifier.  A lower threshold
(e.g., $P > 0.3$) recovers more Type~Ia candidates at the cost of
increased contamination.
\end{enumerate}


%% =========================================================================
\section*{References}
%% =========================================================================

\begin{thebibliography}{9}

\bibitem[Schlegel et al.(1998)]{Schlegel1998}
Schlegel, D.~J., Finkbeiner, D.~P., \& Davis, M.\ 1998,
\textit{ApJ}, 500, 525.
\textit{Maps of Dust Infrared Emission for Use in Estimation of
Reddening and Cosmic Microwave Background Radiation Foregrounds.}

\bibitem[Schlafly \& Finkbeiner(2011)]{Schlafly2011}
Schlafly, E.~F., \& Finkbeiner, D.~P.\ 2011,
\textit{ApJ}, 737, 103.
\textit{Measuring Reddening with Sloan Digital Sky Survey Stellar
Spectra and Recalibrating SFD.}

\bibitem[Villar et al.(2019)]{Villar2019}
Villar, V.~A., et al.\ 2019,
\textit{ApJ}, 884, 83.
\textit{SuperRAENN: A Semi-supervised Supernova Photometric Classification
Pipeline Trained on Pan-STARRS1 Medium-Deep Survey Supernovae.}

\end{thebibliography}


\end{document}
